Equalizer Programming Guide

### 2008-07-08 Developer/ProgrammingGuide/paper.tex
Update docs (namespace renaming)
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.4 for Equalizer 0.5.2+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & File Format Appendix\\
    1.3.4   & July 3, 2008 & Update statistics section\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The distributed data classes are helper classes based on
\textsf{eqNet::Object}. They illustrate the typical usage of distributed
objects for static as well as dynamic, frame-specific data.

@@

The third step is to create an instance of the application and to
initialize it locally. The application is an \textsf{eq::Client}, which
in turn is an \textsf{eqNet::Node}. The underlying network layer in
Equalizer is a peer-to-peer network of \textsf{eqNet::Node}s. The
application programmer does not usually have to be aware of the classes
in the \textsf{eqNet} namespace, but both the \textsf{eq::Client} and
the server are \textsf{eqNet::Node}s.

@@

{\footnotesize\begin{lstlisting}
    // 3. init config
    eqBase::Clock clock;

@@

Equalizer provides distributed objects which help implementing data
distribution in a cluster environment. The master version of a distributed
object is registered with a \textsf{eqNet::Session}, which assigns a
session-unique identifier to the object. Other nodes can map their
instance of the object to this identifier, thus synchronizing the
object's data with the remotely registered master version.

Distributed objects are created by subclassing from
\textsf{eqNet::Object}. Distributed objects can be static (immutable) or
dynamic. Dynamic objects are versioned.

@@

{\footnotesize\begin{lstlisting}
void InitData::getInstanceData( eqNet::DataOStream& os )
{
    os << _frameDataID << _modelID << _windowSystem << _useVBOs << _useGLSL
       << _invFaces ;
}

void InitData::applyInstanceData( eqNet::DataIStream& is )
{
    is >> _frameDataID >> _modelID >> _windowSystem >> _useVBOs >> _useGLSL
       >> _invFaces;

@@

Versioned objects have to override \textsf{getChangeType} to indicate
how they want to have changes to be handled. The current
implementation has the following characteristics:
\begin{itemize}
\item Only the master instance of the object is writable, that is,
  \textsf{eqNet::Object::com\-mit} can be called only on the master
  instance to generate a new version.
\item Slave instance versions can only be advanced, that is,
  \textsf{eqNet::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instance are mapped to the oldest available
  version.
\end{itemize}

@@

{\footnotesize\begin{lstlisting}
void VertexBufferDist::getInstanceData( eqNet::DataOStream& os )
{
    EQASSERT( _node );
    os << _isRoot;

@@

The \textsf{eqPly::Config} class holds the master versions of the
initialization and frame data. Both objects are registered with the
\textsf{eq::Config}, which is the \textsf{eqNet::Session} used for
rendering. Equalizer takes care of the session setup and exit in
\textsf{Client::choose\-Config} and \textsf{Client::releaseConfig},
respectively.

@@

{\footnotesize\begin{lstlisting}
mesh::VertexBufferRoot* VertexBufferDist::mapModel(eqNet::Session* session,
                                                   const uint32_t modelID )
{
    VertexBufferDist modelDist( 0, 0 );

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\small\label{fThreads}Threads within one node process}}
\end{wrapfigure}
The command thread processes Equalizer-related request from other nodes,
for example during \textsf{eqNet::Object} mapping. In some special cases
the command thread executes application code, for example when a remote
node maps a static or unbuffered object,
\textsf{Object::getInstanceData} is called from the command thread.

@@

The file format represents an ascii deserialization of the
server. Streaming an \textsf{eqs::Server} to an \textsf{eqBase::Log}
ostream produces a valid configuration file. Likewise, loading a
configuration file produces an \textsf{eqs::Server}.

@@

The configuration file consist of an optional global section and a
server configuration. The global section defines default values for
various attributes. The server section represents an
\textsf{eqs::Server}.

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.5 for Equalizer 0.5.3+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & File Format Appendix\\
    1.3.4   & July 3, 2008 & Update statistics section\\
    1.3.4   & July 7, 2008 & Update after namespace naming cleanup\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The distributed data classes are helper classes based on
\textsf{eq::net::Object}. They illustrate the typical usage of distributed
objects for static as well as dynamic, frame-specific data.

@@

The third step is to create an instance of the application and to
initialize it locally. The application is an \textsf{eq::Client}, which
in turn is an \textsf{eq::net::Node}. The underlying network layer in
Equalizer is a peer-to-peer network of \textsf{eq::net::Node}s. The
application programmer does not usually have to be aware of the classes
in the \textsf{eq::net} namespace, but both the \textsf{eq::Client} and
the server are \textsf{eq::net::Node}s.

@@

{\footnotesize\begin{lstlisting}
    // 3. init config
    eq::base::Clock clock;

@@

Equalizer provides distributed objects which help implementing data
distribution in a cluster environment. The master version of a distributed
object is registered with a \textsf{eq::net::Session}, which assigns a
session-unique identifier to the object. Other nodes can map their
instance of the object to this identifier, thus synchronizing the
object's data with the remotely registered master version.

Distributed objects are created by subclassing from
\textsf{eq::net::Object}. Distributed objects can be static (immutable) or
dynamic. Dynamic objects are versioned.

@@

{\footnotesize\begin{lstlisting}
void InitData::getInstanceData( eq::net::DataOStream& os )
{
    os << _frameDataID << _modelID << _windowSystem << _useVBOs << _useGLSL
       << _invFaces ;
}

void InitData::applyInstanceData( eq::net::DataIStream& is )
{
    is >> _frameDataID >> _modelID >> _windowSystem >> _useVBOs >> _useGLSL
       >> _invFaces;

@@

Versioned objects have to override \textsf{getChangeType} to indicate
how they want to have changes to be handled. The current
implementation has the following characteristics:
\begin{itemize}
\item Only the master instance of the object is writable, that is,
  \textsf{eq::net::Object::com\-mit} can be called only on the master
  instance to generate a new version.
\item Slave instance versions can only be advanced, that is,
  \textsf{eq::net::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instance are mapped to the oldest available
  version.
\end{itemize}

@@

{\footnotesize\begin{lstlisting}
void VertexBufferDist::getInstanceData( eq::net::DataOStream& os )
{
    EQASSERT( _node );
    os << _isRoot;

@@

The \textsf{eqPly::Config} class holds the master versions of the
initialization and frame data. Both objects are registered with the
\textsf{eq::Config}, which is the \textsf{eq::net::Session} used for
rendering. Equalizer takes care of the session setup and exit in
\textsf{Client::choose\-Config} and \textsf{Client::releaseConfig},
respectively.

@@

{\footnotesize\begin{lstlisting}
mesh::VertexBufferRoot* VertexBufferDist::mapModel(eq::net::Session* session,
                                                   const uint32_t modelID )
{
    VertexBufferDist modelDist( 0, 0 );

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\small\label{fThreads}Threads within one node process}}
\end{wrapfigure}
The command thread processes Equalizer-related request from other nodes,
for example during \textsf{eq::net::Object} mapping. In some special cases
the command thread executes application code, for example when a remote
node maps a static or unbuffered object,
\textsf{Object::getInstanceData} is called from the command thread.

@@

The file format represents an ascii deserialization of the
server. Streaming an \textsf{eq::server::Server} to an \textsf{eq::base::Log}
ostream produces a valid configuration file. Likewise, loading a
configuration file produces an \textsf{eq::server::Server}.

@@

The configuration file consist of an optional global section and a
server configuration. The global section defines default values for
various attributes. The server section represents an
\textsf{eq::server::Server}.
